Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention

### 2024-06-28 manuscripts/manuscript1_supplement.qmd
prior predictive check function

library(targets)
library(tidyverse)
library(withr)
library(here)
library(modelsummary)
library(marginaleffects)
library(brms)
library(tidybayes)
library(bayesplot)
library(glue)
library(ggdist)
library(patchwork)
library(tinytable)
library(flextable)

@@

# visualization function
results_visual <- function(model, title) {
  model <- model
  subtitle <- title

  draws <- model |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize()

  pred <- predictions(model, by = "ich_laterality") |>
    tibble()

  theme_ich <- function(base_family = "Inter",
                        base_size = 14) {
    theme_minimal(base_size = base_size) +
      theme(
        text = element_text(lineheight = 2),
        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
        axis.text.y = element_text(size = rel(0.8)),
        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
        axis.text.x = element_text(size = rel(2)),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.justification = 1,
        panel.grid = element_line(colour = "#F3F4F5"),
        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
      )
  }

  p1 <- predictions(model, by = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw, fill = ich_laterality)) +
    geom_density(alpha = .7) +
    scale_fill_manual(values = c("#422f05", "#E69F00")) +
    scale_x_continuous(labels = scales::percent) +
    labs(
      title = "Predicted Probability",
      fill = "ICH Laterality",
      y = NULL
    ) +
    theme_ich()

  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw)) +
    geom_density(color = "#422f05", fill = "#E69F00") +
    scale_x_continuous(labels = scales::percent) +
    labs(
      title = "Average Difference between Left vs. Right Hemisphere",
      y = NULL
    ) +
    geom_vline(
      xintercept = 0, linetype = "dashed",
      color = "#999999", linewidth = 1
    ) +
    theme_ich()

  p3 <- model |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    ggplot(aes(ich_right_or)) +
    geom_density(color = "#422f05", fill = "#E69F00") +
    # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
    scale_x_continuous(breaks = seq(0, 3, 0.25)) +
    labs(
      title = "Odds Ratio - Right Hemisphere",
      y = NULL,
      fill = NULL
    ) +
    geom_vline(
      xintercept = 1, linetype = "dashed",
      color = "#999999", linewidth = 1
    ) +
    theme_ich() +
    theme(legend.position = "none")

  total <- p1 / p2 / p3

  final <- total +
    plot_annotation(
      title = glue(subtitle),
      theme = theme(plot.title = element_text(size = rel(2), margin = margin(12, 0, 8, 0)))
    )

  return(final)
}

@@

# Return dnorm(x) for 0 < x < 2, and NA for all other x
dnorm_limit <- function(x) {
  y <- dnorm(x)
  y[x < 0] <- NA
  return(y)
}

ggplot(data = tibble(x = -2:2), aes(x = x)) +
  geom_function(fun = dnorm, args = list(mean = -20, sd = 0.5), linewidth = 1) +
  # stat_function(fun = dnorm_limit, geom = "area", color = "orange", alpha = 0.2) +
  geom_vline(xintercept = -0.22, linetype = "dashed", color = "orange", linewidth = 1) +
  scale_x_continuous(breaks = seq(-2, 2, 0.5)) +
  theme_ich()

@@

```{r informative prior predictive check - neurosurgical intervention}
#| echo: false
#| fig-height: 12
#|

with_dir(here(), tar_load(m_prior_neutral_neurosurgery))
with_dir(here(), tar_load(m_prior_left_neurosurgery))

avg_predictions(m_prior_left_neurosurgery, variables = "ich_laterality")

results_visual(m_prior_left_neurosurgery, "Prior Predictive Check")

>>>>

#| echo: false
#| fig-height: 12

library(targets)
library(tidyverse)
library(withr)
library(here)
library(modelsummary)
library(marginaleffects)
library(brms)
library(tidybayes)
library(bayesplot)
library(glue)
library(ggdist)
library(patchwork)
library(tinytable)
library(gt)

@@

@@

# ggplot(data = tibble(x = -2:2), aes(x = x)) +
#   geom_function(fun = dnorm, args = list(mean = -20, sd = 0.5), linewidth = 1) +
#   # stat_function(fun = dnorm_limit, geom = "area", color = "orange", alpha = 0.2) +
#   geom_vline(xintercept = -0.22, linetype = "dashed", color = "orange", linewidth = 1) +
#   scale_x_continuous(breaks = seq(-2, 2, 0.5)) +
#   theme_ich()

```{r prior predictive check - function}
#| echo: false
#| output: false
## | # | | fig-height: 12

f_prior_predictive_check <- function(neutral_model, left_model, right_model, flat_model) {
  models <- list(
    "Neutral Prior" = neutral_model,
    "Left Hemisphere Prior" = left_model,
    "Right Hemisphere Prior" = right_model,
    "Flat Prior" = flat_model
  )

  # gather summary variables for each model
  neutral <- models$"Neutral Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  left <- models$"Left Hemisphere Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  right <- models$"Right Hemisphere Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  flat <- models$"Flat Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  prior_summary <- rbind(neutral, left, right, flat) |>
    mutate(prior = c("Neutral", "Left Hemisphere", "Right Hemisphere", "Flat")) |>
    select(c(prior, aor, lower_ci, upper_ci, or_1, rope))

  # add prior predictions
  neutral_predictions <-
    as_tibble(avg_predictions(models$"Neutral Prior", by = "ich_laterality")) |>
    mutate("prior" = "Neutral") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  left_predictions <-
    as_tibble(avg_predictions(models$"Left Hemisphere Prior", by = "ich_laterality")) |>
    mutate("prior" = "Left Hemisphere") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  right_predictions <-
    as_tibble(avg_predictions(models$"Right Hemisphere Prior", by = "ich_laterality")) |>
    mutate("prior" = "Right Hemisphere") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  flat_predictions <-
    as_tibble(avg_predictions(models$"Flat Prior", by = "ich_laterality")) |>
    mutate("prior" = "Flat") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  prediction_summary <- rbind(neutral_predictions, left_predictions, right_predictions, flat_predictions)

  prior_prediction_summary <- left_join(prediction_summary, prior_summary, by = "prior") |>
    gt(rowname_col = "prior") |>
    tab_stubhead(label = "Type of Prior") |>
    fmt_number(columns = everything(), decimals = 2) |>
    fmt_number(columns = 6, rows = 4, decimals = 0) |>
    fmt_percent(columns = 2:3, rows = 1:4, decimals = 1) |>
    tab_header(
      title = md("**Prior Predictive Checks**")
    ) |>
    cols_label(
      Left = "Left Hemisphere",
      Right = "Right Hemisphere",
      aor = "aOR",
      lower_ci = "Lower 95% Credible Interval",
      upper_ci = "Upper 95% Credible Interval",
      or_1 = "Probability of Difference (aOR >1)",
      rope = "% of Distribution outside of ROPE"
    ) |>
    cols_width(
      1 ~ pct(50)
    )

  return(prior_prediction_summary)
}

@@

```{r prior predictive check - neurosurgery}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_neurosurgery))
left_model <- with_dir(here(), tar_read(m_prior_left_neurosurgery))
right_model <- with_dir(here(), tar_read(m_prior_right_neurosurgery))
flat_model <- with_dir(here(), tar_read(m_prior_flat_neurosurgery))

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

### Ventriculostomy

```{r prior predictive check - evd}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_evd))
left_model <- with_dir(here(), tar_read(m_prior_left_evd))
right_model <- with_dir(here(), tar_read(m_prior_right_evd))
flat_model <- with_dir(here(), tar_read(m_prior_flat_evd))

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

predictions(neutral_model, by = "ich_laterality") |>
  posterior_draws() |>
  ggplot(aes(x = draw, fill = ich_laterality)) +
  geom_density(alpha = .3) +
  scale_fill_manual(values = c("#422f05", "#E69F00")) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    title = "Predicted Probability",
    fill = "ICH Laterality",
    y = NULL
  )

avg_comparisons(neutral_model, variables = "ich_laterality") |>
  posterior_draws() |>
  ggplot(aes(x = draw)) +
  geom_density(color = "#422f05", fill = "#E69F00") +
  scale_x_continuous(labels = scales::percent) +
  labs(
    title = "Average Difference between Left vs. Right Hemisphere",
    y = NULL
  ) +
  geom_vline(
    xintercept = 0, linetype = "dashed",
    color = "#999999", linewidth = 1
  )

neutral_model |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  ggplot(aes(ich_right_or)) +
  geom_density(color = "#422f05", fill = "#E69F00") +
  # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
  scale_x_continuous(breaks = seq(0, 3, 0.25)) +
  labs(
    title = "Odds Ratio - Right Hemisphere",
    y = NULL,
    fill = NULL
  ) +
  geom_vline(
    xintercept = 1, linetype = "dashed",
    color = "#999999", linewidth = 1
  )
